---
title: "Code - Note 5"
subtitle: "Method of Moment estimation for models on block graphs"
author: "Stefka Asenova"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Code - Note 5}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: mypr1_rev1.bib
---

<style type="text/css">
  body{
  font-size: 12pt;
  max-width: 1000px;
  margin-left: auto;
  margin-right: auto;
}
</style>



```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(gremes)

```

## Generate the data first

We do not have a real dataset according to a given block graph which is not a tree. Therefore we first generate the observations. How is the data generated is described in Vignette "Additional functionalities".

Create the block graph.

```{r fig.height=4.5, fig.width=4.5}
g<- graph(c(1,3,1,2,2,3,
            3,4,4,5,5,3,
            3,7,3,6,6,7), directed=FALSE)
g<- set.vertex.attribute(g, "name", V(g), c("a", "b", "c", "d", "e", "f", "g"))
plot(g)

```

Create the edge weights to be assigned to the edges of the graph.

```{r}
# all deltas are squares already
C1<- c(0.2, 0.8, 0.6)     # d_13^2, d_12^2, d_23^2
C2<- c(0.3, 0.5, 0.1)     # d_34^2, d_45^2, d_35^2
C3<- c(0.4, 0.05, 0.25)   # d_37^2, d_36^2, d_67^2

```

Attach the edge weights to the edges.

```{r}
hrmbgobj<- HRMBG(g)
hrmbgobj<- setParams(hrmbgobj, c(C1, C2, C3))
hrmbgobj

```

Create the matrix $\Lambda$, whose entry $\lambda_{ij}$ is the sum of the edge weights on the unique shortest path between node $i$ and node $j$.

```{r}
hrmlam<- HRLambda(hrmbgobj)
hrmlam

```

Generate 1000 observations from Huesler-Reiss distribution with parameter matrix $\Lambda$ and some independent
random noise.

```{r}
X<- rHRM(hrmbgobj, hrmlam, 1000, noise = TRUE)

```

## Perform estimation 

Create the subsets for local estimation.

```{r}
rdsobj<- RootDepSet()
rdsobj<- setRootDepSet(rdsobj, subset = list(c("a", "b", "d", "e", "f", "g"),
                                             c("a", "b", "d", "e", "f", "g"),
                                             c("a", "b", "d", "e", "f", "g"),
                                             c("a", "b", "d", "e", "f", "g"),
                                             c("a", "b", "d", "e", "f", "g"),
                                             c("a", "b", "d", "e", "f", "g")), 
                       c("a", "b", "d", "e", "f", "g"))

```

Estimate the model treating the third variable as latent: create first an object of class `HRMBG` and then use on it the method `estimate`.

```{r}
hrmbg<- HRMBG(g)
hrmbg<- suppressMessages(estimate(hrmbg, X[,-3], rdsobj, k_ratio=0.2))
hrmbg$depParams

```

We have suppressed the messages. The messages are informative, if they don't stop the process they are not errors.


